__________________________
Homework 4: Topics 10 & 11
__________________________


1) What is the purpose of a linear regression? When would you use it in your research?


Linear regression is used to model and predict a variable given one or more predictor variables to do so. You should use it when you want to see if your predictor variables have a relationship with the predicted variable.
_______________________________________________________________________________________



2)What should you do with a categorical variable in a regression?

When dealing with a categorical variable you should use the as.factor() method so that they can be properly included in the regression


_______________________________________________________________________________________


In this section, run the regression model and answer the following questions

```{r}
setwd("C:/R")
study_meditation <- read.csv("Study_meditation.csv")

```

```{r}
study_gender <- lm(Score ~ Study.Hours+Gender_Dummy, data = study_meditation)

summary(study_gender)

#in the Gender_Dummy variable, female = 1, male = 0

```

3) What is the base score?

___________
The base score is around 52.8
___________

4) How many points are earned per hour of study?

___________
around 6.3 points are earned per hour of studying.
___________

5) How many points are earned based on gender?

__________
The model predicts that women will score 0.93 points higher than their male counterparts
__________

6) What would the predicted score be for a male student that studied for 2 hours? Show your work

```{r}
male_score <- 52.8336 + 6.3024 * 2
male_score
```
_________

_________

7) What would the predicted score be for a female student that studied for 3 hours? Show your work

```{r}
female_score <- 52.8336 + (6.3024 * 3) + 0.9306
female_score
```
_________

_________

8) I think that class also has an effect on score. Using the following model, what is the predicted score for a male student in class 3 that studied for 6 hours. Show your work

```{r}
library(dplyr)

study_mutate <- study_meditation %>% 
  mutate(Class = as.factor(Class))

study_gender_class <- lm(Score ~ Study.Hours+Gender_Dummy+Class, data = study_mutate)

summary(study_gender_class)
```

```{r}
class_score <- 44.454 + (7.128 * 6) + 2.660
class_score
```
____________

____________

9) Using the same model from question 8, what is the predicted score for a female student in class 2 that studied for 4 hours. Show your work

```{r}
female_class <- 44.454 + (7.128 * (4)) - 1.224 + 15.522
female_class
```
__________


__________

10) Lastly, I think meditation hours impacts each class differently. What would be the predicted score of a male student in class 3 that studied for 3 hours with 2 hours of meditation? Show your work

```{r}
study_medhours <- lm(Score ~ Study.Hours+Gender_Dummy+Class*Meditation_Hours, data = study_mutate)

summary(study_medhours)
```
```{r}
meditation <- 107.8066 + (4.8327 * 3) - 49.7770 - (49.4551 * 2) + (48.0827 * 2)
meditation
```
______


______
