---
title: "HW: 4 Topic 10, Linear Regression"

---

```{r}
library(dplyr)
library(car)
library(lsr)
library(ggplot2)
options(scipen=999)
```

Data Source: https://howmuch.net/articles/hours-work-afford-home

```{r}
setwd("/Users/aidenstanton/projects/R")
housing <- read.csv("Housing.csv")
```

In this example we want to build a regression model that shows the relationship between the Hours.per.Month.to.Afford.a.Home and  and Median.Home.Listing.Price. We predict that the median home price influences the hours of work needed to afford a home. Median home price is our independent variable and Hours per month is our dependent variable. 

1) Let's set up our regression and create a summary output

```{r}
housing_regression <- lm(formula = Hours.per.Month.to.Afford.a.Home ~ Median.Home.Listing.Price, data=housing)
summary(housing_regression)
```

2) Explain each part of the coefficents and what it means for our regression 
_________________________________________________________________
The intercept it the hours when price should be = 0. 
Median home listing price = .000087812 for every $1 increase in price have to work that many hours. Higher prices work longer.
Std error shows that the slope in reliable
small p value means statistically significant

_________________________________________________________________

3) Plot the regression

```{r}
ggplot(housing, aes(x = Median.Home.Listing.Price, y = Hours.per.Month.to.Afford.a.Home))+
  geom_point()+
  geom_smooth(method = "lm", se = F)+
  theme_minimal()

```
4) Let's see if another predictor variable will change our model. I also think that the Monthly.Mortgage.Payment influences Hours.per.Month.to.Afford.a.Home. Create a new regression with both of the predictor variables. Include the summary output.

```{r}
housing_regression2 <- lm(formula = Hours.per.Month.to.Afford.a.Home ~ Median.Home.Listing.Price +  Monthly.Mortgage.Payment, data=housing)
summary(housing_regression2)

```
5)  Interpret what the coeffiecents in the output is telling us about the regression model.
_____________________________________________________________________________
Intercept is the same as before
Median home listing price has an insignificant p value with mortgage payments not useful as a predictor anymore.
Monthly mortage payment also has an insignificant p value and also has a negative coefficient so higher mortgage payments decrease hours.

_____________________________________________________________________________

6) Compare the regression models. Which model seems like a better fit for our data and why?
The first model is better because it has a significant predictor. 


7) Based on this analysis, what inferences could you make about housing affordability and work hours?

Based on this analysis we can infer that housing affordability is affected by home prices as prices increase people have to work more hours to afford a home which can make it hard to keep up with ever increasing costs.