---
title: "Topic 7: Hypothesis Testing"
output: html_notebook
---

So, thus far we've talked a lot about hypothesis testing in theory. Now it's time to see what it looks like in practice, using R!

```{r}
#Null: The probability of flipping heads is 50%, or θ = 0.5
#Alternative: The probability of flipping heads is not 50%, or θ ≠ 0.5 (this is what we predict)

#We can replicate the hypothesis test from the lecture very easily with the binom.test() function:
#Remember that we are using a binomial distribution because out data only has two outcomes: heads or tails. The number of participants who get heads on a coin flip is between 0 and 100.

#in our experiment, 70 people land on heads out of 100 participants.
options(scipen = 999)
#I'm using scipen here to turn off scientific notation
binom.test(x=70, n=100, p=.5 )

#x=70 because that was the result of our experiment (X = the observed "success")
#n=100 is the number of participants
#p = .5 is the probability of flipping a head (p would equal the probability of a "successful" outcome under the null hypothesis)
```


I know this output is a little intimidating! Don't worry, you'll get used to running these kinds of tests as we go along. With a p-value of 0.0000785(a very small number!) we can reasonably reject the null hypothesis and conclude the probability of flipping a head with this coin is not equal to 0.5. 

```{r}
#What if we only want to know if our sample is greater than 50? We can perform a one-sided hypothesis test like this:

binom.test(x=70, n=100, p=.5, alternative = "greater")

```
*using the confidence interval is concluding that the true probability of success at the population level is somewhere between .615 and 1

Using the 1 sided hypothesis test, we can also reject the null hypothesis.  


This week, you learned some of the basic theory and assumptions behind a hypothesis test. 

-----------------------------------------------------------------------------
In class activity: Let's think about setting up some hypothesis tests and running them. 
-----------------------------------------------------------------------------

Group 1: I am rolling one six-sided die 50 times. I roll a 3 40/50 times. Design a test to show whether or not the die is fair and report your results.


1) What is the alternate hypothesis and what is the null hypothesis?

 null-the die is fair and the probability of rolling a 3 is 1/6
 alternate- the die is not fair and the probability of rolling a 3 is not 1/6

2) Design the test to show your results

```{r}
binom.test(x=40, n=50, p=(1/6))
```

3)Using a significance level of .05, Can we reject the null hypothesis

Yes.  The p-value is way under .001, the lowest of the standard alphas. In our results, with 95% confidence, the probability of rolling a 3 with this die is between .66-.9 (which is really high and gives us a broader range of possible estimates for a larger population). Our sample estimate probability of success is .8

________________________________
Group 2: I have a 10-sided die, and I role it 100 times. I roll a 2 20 times. Design a test to show whether the die is fair. 

4) What is the alternate hypothesis and what is the null hypothesis?
 
 null-the die is fair and the probability of rolling a 2 is 1/10
 alternate- the die is not fair and the probability of rolling a 2 is not 1/10

5) Design the test to show your results

```{r}
binom.test(x=20, n=100, p=(1/10))
```

6) If we have an alpha=.001, Can we reject the null hypothesis

No  not at .001 because our p-value is higher than .001. We could reject at an alpha of .05 and .01, 

______________________________________________________

Group 3: I have a bag of marbles: 3 blue, 2 green, and 1 red. I draw from the bag (with replacement) 50 times. I drew a blue marble 35 times. Design a test to show whether we pulled a blue marble MORE than 50% of the time.


7) What are the alternate and null hypotheses
Alt: The probability of drawing a blue marble is > .5
Null: The probability of drawing a blue marble is a blue marble is .5 or 50%


8) Design a binom Test

```{r}
#Answer here

binom.test(x=35, n=50 , p=.5, alternative = "greater")

```
9) At which significance levels can we reject the null hypothesis? 

yes at the .05 and .01 alpha
______________________________________________________

Group 4: I am running a study to test for evidence of psychic ability among my participants. A lab technician holds up a card so that the back is facing the participant and the front of the card is facing the technician. Participants need to guess if the card has an image on it or if it is blank.  Out of 900 participants 467 are correct with their guess. 

10) What are the alternate and null hypotheses
Alt: The probability of guessing correctly is greater than .5
Null: The probability guessing correctly is .5

11) Design a binom test
```{r}
binom.test(x=467, n=900, p=.5, alternative = "greater")
```
12) Are our participants significantly more psychic than what we would predict about the larger population?
No, our p value is not less than .05

