---
title: "Week 6: Confidence Intervals"
output: html_notebook
---

This week, most of the lesson was in lecture format. We'll just go over a quick tutorial on how to calculate confidence intervals in R. 

Confidence Intervals: A range of plausible values for an unknown population parameter, like mean, calculated from your sample. We almost never know the parameters for any given population and can use our sample to help generate estimates. Confidence intervals refer to the reliability of the sampling method. A confidence interval of 75% would mean that if the sampling process was repeated multiple times, 75% of those intervals would contain the true population parameter. The higher our confidence interval, the more precise our estimate is. 


As I said in lecture, we'll use the t-distribution to calculate confidence intervals, rather than the normal distribution. 

Rember, the T-Distribution is

-smooth like the normal distribution
-symmetric around the mean
-has a mean of 0
-Unlike a normal distribution, this does not require us to assume that the population standard deviation (SD) is known
-It is better suited for smaller samples (e.g. usually 30 or below); as the sample size increases, it approaches the normal distribution 

```{r}
#In practice, there's a very simple function in R that will calculate all of this for you (below), but knowing how to use the T-distribution is helpful information. 

#We calculate the 2.5th and 97.5th percentiles of the normal distribution like this. The qnorm gives us an sd of +- 1.96 sd away from the mean and it captures 95% of the central distribution (this is the middle most of the data). 

#we use qnorm  because calculates quantiles                

qnorm( p = c(.025, .975) )

#This is where the 1.96 comes from in the formula!
```
```{r}
#We can do the same with the T-distribution! The T-distribution does not require us to assume that the population SD is known. We just need the sample size to compute degrees of freedom. I;ll start with making an object called "N" for the size of our sample.

N <- 10000   # suppose our sample size is 10,000
#we assign our sample size to an object named "N" (replicating the formula from the presentation)
qt( p = .975, df = N-1)
#N= our sample size
#qt= quantile t-distribution function
#p=percentile
#df= degrees of freedom = sample size-1 (N-1)

```
Here you can see that, with a large sample size, the T-distribution is roughly equivalent to the normal distribution. Typically, T-distributions are used with sample sizes under 30. Let's take a look at a smaller sample size. 

```{r}
N <- 10   # suppose our sample size is 10
qt( p = .975, df = N-1)
```
How do we use this information? We would plug it into the following formula:

                        x-bar +/- (critical value * sigma-hat)
                        
        sample mean +/- (call to qt * estimate of population sd using N-1 in denominator)
                                                                      
Where the result of the call to qt() (in this example ~2.26) is the critical value, x-bar is our sample mean, and sigma-hat is our estimate of the population standard deviation (using N-1 in the denominator). 

```{r}
#While there is no built-in base R function to calculate confidence intervals, we can use the ciMean() function from the LSR package to calculate CIs for us

library(lsr)
library(dplyr)

#Let's load back in the temperature data

setwd("/Users/aidenstanton/projects/R")
temps <- read.csv("temps_nyc.csv")

temps <- read.csv("temps_nyc.csv")

#create month and year variables and select only 2014
#the original data frame spans from July of 2014-June of 2015

temps_14 <- temps %>% 
  mutate(month = format(as.Date(date, format = "%m/%d/%Y"),"%m"),
         year = format(as.Date(date, format = "%m/%d/%Y"), "%Y")) %>% 
  filter(year == "2014")

#new dataframe <- created from original data frame then(pipe) mutate (add columns month, year) 
#as.Date (date, format = "%m/%d/%Y"), "%m" Converts the date column from the original dataframe (temps) into a Date object. format= "%m/%d/%Y" tells R how to interpret the dates in the date column from the original dataframe(temps). In temps, the date is written out as "mm/dd/YY"  The same format is used for the year column, except instead of the column containing the month, it will state the year.  Then we are telling R to filter the dataframe by 2014. 

#________________________________________________

#Let's calculate the 95% confidence interval
ciMean(temps$actual_mean_temp)

#ciMean is calculating the average of the variable actual_mean_temp (or the data that you're looking at). 
#then ciMean is caclulating the range around the mean, using the standard deviation. This is defaulted to using a confidence interval of 95%. This would mean that there is a 95% probability that the confidence interval  will contain the true population mean. 

#the output is telling us the upper and lower bounds of the 95% confidence interval.  With a confidence interval of 95%, the true mean of the population (which is never truly known) is somewhere between 52.8 and 56.7 degrees
```
```{r}

#We can use the conf =  argument to change the interval that we calculate
ciMean(temps$actual_mean_temp, conf = 0.80)

#How do we interpret? 
```
#With a confidence interval of 80%, the true mean of the population is somewhere between 53.5 and 55.99 degrees
```{r}

#We can plot a confidence interval using the sciplot package - download it now

library(sciplot)
library(lsr)

#we'll start with a bargraph - you can plot it using the code below. We need a grouping variable so that we can plot confidence intervals across groups (i.e. months, years, etc.)

#If we only need to calculate one confidence interval, then there's really no need to graph it! 

bargraph.CI(  x.factor = month,                # Grouping variable
              response = actual_mean_temp,     # outcome variable
              data = temps_14,                 # data frame with the variables
              ci.fun= ciMean,                  # name of the function to calculate CIs
              xlab = "Month",                  # x-axis label
              ylab = "Average Temperature"     # y-axis label
 )
#the whiskers the (the lines that look like this |-| but vertical) represent the confidence interval. The upper bar represents the upper bound and the lower bar represents the lower bound. 

#with 95% confidence, the true mean of the temp in the month of November is between 43 and 50 degrees
```

How do we interpret this? The bar shows the mean temperature for each month in 2014. The lines are the 95% confidence intervals. 

```{r}
#We can also plot confidence intervals using line graphs, like this:

lineplot.CI( x.factor = month,                 # grouping variable 
             response = actual_mean_temp,      # outcome variable
             data = temps_14,                  # data frame with the variables
             ci.fun= ciMean,                   # name of the function to calculate CIs
             xlab = "Month",                   # x-axis label
             ylab = "Average Temperature "     # y-axis label
)
```



And there you have it, short and sweet! We'll come back to more R next week for hypothesis testing. 
