---
Topic 7

---

```{r}
library(lsr)
library(dplyr)
```


____________
Topic 7
____________

1) What is a p-value? Why can't you just report the p-value of your data/variable?

A number from 0 to 1 indicating the probability of getting a specific result
_________________________________________________________________________________________________
2) What is the difference between type 1 and type 2 error?

Type 1 errors are seen as bigger errors occur when we reject a value that is actually true while a type 2 is rejecting a value while the alternate hypothesis is true

__________________________________________________________________________________________________

3) At which alpha/critical regions is this p-value significant? 

0.05
0.01
0.001
____________________________________________________________________________________________________

I am rolling one 12-sided die 18 times. I roll an 11  6/18 times. Design a test to show whether or not the die is fair and report your results.


1) What is the alternate hypothesis and what is the null hypothesis?
Null hypothesis is that you should roll an 11 every 1/12 times
the alternate hypothesis is that you will roll it more or less than every 1/12 times


2) Design the test to show your results

```{r}
binom.test(6, 18, p = (1/12))
```

3)Using a significance level of .05, Can we reject the null hypothesis
No because our value is higher than .05

__________________________________________________________________________________________________
